---
title: "PE Delivery Predictor Manual"
date: "03/06/2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

The web application provides predictive insights into the time-to-delivery of patients diagnosed with preeclampsia. It uses pre-trained models to provide accurate predictions and visualizations, making it a user-friendly tool for healthcare professionals.


### Step 1: Choose the panel

The web application has two main panels: the individual prediction panel and the group prediction panel.

- Individual prediction panel: This panel calculates the prognosis score (PI) of time to delivery for a single new patient using pre-trained models. The input should be a text file with one line of values from a single patient.
- Group prediction panel: This panel takes a group of new patients as input and returns the predicted PIs and their percentiles in a table format.

### Step 2: Choose pre-trained model

The user needs to choose the desired pre-trained model based on their patient characteristics and data availability. The available pre-trained models are:

```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
|Pre-trained Model   |Description                                |Required Input                                       |
|:--------------------|:--------------------------------------------|:-----------------------------------------------------|
| Baseline Model     | Applicable to all PE patients              |Baseline variables                                   |
| Full Model         | Applicable to all PE patients              |Baseline variables, lab results and vital measuements|
| EOPE Baseline Model| Applicable to only early onset PE patients |Baseline variables                                   | 
| EOPE Full Model    | Applicable to only early onset PE patients |Baseline variables, lab results and vital measuementS| 
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

### Step 3: Upload new patient data
Both the individual prediction panel and the group prediction panel require a single delimited file as input. The users have the option to choose between three delimiters: ";", ",", or ":". The input file must have the same columns in the same order as specified. The individual prediction panel requires only a single row of data in the input file, whereas the group prediction panel allows for multiple rows.


```{r table2, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
|Selected Model       |Required Variables   |Variable Description                                                                              |
|:--------------------|:--------------------|:-------------------------------------------------------------------------------------------------|
| Baseline Model     |diag_GA               |PE diagnoses gestational age(days)                                |
|                    |SeverePE              |1:  the preeclampsia is severe at the first diagnosis; 0: mild preeclampsia at the first diagnosis|
|                    | past_pe              |1:  the patient has previous preeclampsia history; 0:Otherwise                                    |
|                    | age                  |Maternal age at the beginning of current pregnancy (Years)                                 |
|                    | EPIS_PARA_COUNT      |Previous parity counts                                                                     |
|                    | DiabetesUncomplicated|1: The patient has  uncomplicated diabetes comorbidity; 0:Otherwise                                  |
|                    | ValvularDisease      |1: The patient has  valvular disease comorbidity; 0:Otherwise                                    |
| Full Model         |diag_GA               |PE diagnoses gestational age(days)                                |
|                    |SeverePE              |1:  the preeclampsia is severe at the first diagnosis; 0: mild preeclampsia at the first diagnosis|
|                    |AST                   |Aspartate aminotransferase measurement at diagnosis (u/L)                                         |
|                    |RRSD                  |Standard deviation of respirtory rate within 5 days before diagnosis                              |
|                    |EPIS_PARA_COUNT       |Previous parity counts                                                                     |
|                    |creatinine_value      |Creatinine value measurement at diagnosis |
|                    |BPDiaSD               |Standard deviation of diastolic blood pressure within 5 days before diagnosis   |
|                    |age                   |Maternal age at the beginning of current pregnancy (Years)                                 |
|                    |White_Blood_Cell_Count|White blood cell count at diganosis |
|                    |BPDiaMean             |Average of diastolic blood pressure within 5 days before diagnosis |
|                    |Platelet_Count        |Platelet count at diagnosis |
|                    |White_Blood_Cell_Count|White blood cell count at diagnosis|
|                    |past_pe              |1:  the patient has previous preeclampsia history; 0:Otherwise                                    |
|EOPE Baseline Model |diag_GA               |PE diagnoses gestational age(days)                                |
|                    | past_pe              |1:  the patient has previous preeclampsia history; 0:Otherwise                                    |
|                    | EPIS_PARA_COUNT      |Previous parity counts                                                                     |
|                    | Coagulopathy         |1: The patient has Coagulopathy comorbidity; 0:Otherwise                                  |
|                    | PulmonaryCirculationDisorders     |1: The patient has pulmonary circulation disorder; 0:Otherwise                                    |
|                    |SeverePE              |1:  the preeclampsia is severe at the first diagnosis; 0: mild preeclampsia at the first diagnosis|
|EOPE Full Model     | EPIS_PARA_COUNT      |Previous parity counts                                                                     |
|                    | NUMBER_OF_FETUSES    |Baseline variables, lab results and vital measuementS| 
|                    |diag_GA               |PE diagnoses gestational age(days)                                |
|                    | Coagulopathy         |1: The patient has Coagulopathy comorbidity; 0:Otherwise                                  |
|                    | PulmonaryCirculationDisorders     |1: The patient has pulmonary circulation disorder; 0:Otherwise                                    |
|                    |SeverePE              |1:  the preeclampsia is severe at the first diagnosis; 0: mild preeclampsia at the first diagnosis|
|                    |creatinine_value      |Creatinine value measurement at diagnosis|
|                    |Platelet_Count        |Platelet count at diagnosis|
|                    |AST                   |Aspartate aminotransferase measurement at diagnosis (u/L)                                      |
|                    |BPDiaMean             |Average of diastolic blood pressure within 5 days before diagnosis|
|                    |BPSysMean             |Average of systolic blood pressure within 5 days before diagnosis|
|                    |RRSD                  |Standard deviation of respirtory rate within 5 days before diagnosis  |
|                    | past_pe              |1:  the patient has previous preeclampsia history; 0:Otherwise                                    |
"
cat(tabl)
```


### Step 4: Choose whether to scale new data

The pre-trained model mandates the scaling of all numeric variables by dividing them by the square root of their mean value. The variables "creatinine_value," "AST," and "RRSD" must be log-transformed prior to scaling. If the option "Does the data need to be processed?" is set to "Yes," the app will automatically scale the input data. However, if set to "No," it will be the responsibility of the user to scale the data, which is not recommended.

